---
title: "NAMCS Ambulatory Care Services Utilization Among Cancer Patients"
author: "Abhi VIshnu"
date: "2024-01-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = FALSE)
```

## Abstract
Health care costs have been rising at an alarming rate in the USA. The cost of insurance premiums have been increasing, as well as the cost to the CMMS which manages the Medicare programs of millions of Americans. One way to lower the rising costs is by providing preventive services to patients who may be high risk for developing healthcare conditions/complications in the future.

The Center for Disease Control and Prevention conducts a National Ambulatory Medical Care Survey among carefully chosen Physicians in the US to understand services provided to patients in physician offices. A very small number of physicians (a few hundreds) are carefully selected to represent all the physicians in the US, including representation by region, demographics and medical specialty. As this is a complex survey, analysis of data from this survey must be done after applying survey weights as suggested by the CDC.  


# Steps
Here we are going to use the data from the 2015-2018 surveys of the NAMCS. The survey was completed and data released twice in this study period - in 2016 and in 2018.
We used the publicly available data from the CDC website for this analysis.

First, we have to load the required libraries

```{r}
library(haven)
library(naniar)
library(ggplot2)
library(dplyr)
```

Reading datasetss: I have already downloaded the datasets from the CDC website https://www.cdc.gov/nchs/ahcd/datasets_documentation_related.htm
Note that the data is available to download only in SAS, Stata or SPSS formats. To use them in R, we first need to convert them to a R dataset. Here, I have used the *haven* library to read the SAS dataset into R


```{r pressure}
namcs2016_sas <- read_sas("C:/Users/Abhi/Documents/NAMCS/namcs2016_sas/namcs2016_sas.sas7bdat", NULL)
namcs2018_sas <- read_sas("C:/Users/Abhi/Documents/NAMCS/namcs2018_sas/namcs2018_sas.sas7bdat", NULL)

```

We plan to combine the 2016 and 2018 datasets but I noticed that some of the variables are not common. We thus remove those variables to allow us to merge the two datasets.
```{r}
namcs2016_sas_1 <- subset(namcs2016_sas, select = c(-SPECR, -REGIONOFF))
namcs2018_sas_1 <- subset(namcs2018_sas, select = c(-CATPROC1, -CATPROC2, -CATPROC3, -CATPROC4, -CATPROC5, -CATPROC6, -CATPROC7, -CATPROC8, -CATPROC9))
comb201618_namcs <- rbind(namcs2016_sas_1, namcs2018_sas_1)
dim(comb201618_namcs)
```




This dataset has 23118 rows (1 row for each patient) and 1034 columns (variables). Also, each of these patients have been assigned a weight generated using sophisticated complex survey methodology. Thus, all statistics performed with dataset should use those weights to generate estimates that are representative of the whole nation 
#need to replace '-9' '-998' with NA

```{r}
comb201618_namcs <- comb201618_namcs %>% naniar::replace_with_na(replace = list(BPDIAS = c(-9, 998), 
                                                                    BPSYS = c(-9,998),
                                                                    TGSRES = c(-9, 998,-7),
                                                                    LDLRES = c(-9, 998,-7),
                                                                    A1CRES = c(-9, 998,-7),
                                                                    FBGRES = c(-9, 998,-7),
                                                                    PASTVIS = c(-7),
                                                                    PAYTYPER = c(-9, -8),
                                                                    BIOPROV = c(-9, -7),
                                                                    SIGPROV = c(-9, -7),
                                                                    EXCIPROV = c(-9, -7),
                                                                    TOTCHRON = c(-9)))
```

Here we are going to create two summary variables: (1) *TotalServices* to calculate the number of services provided to individual patient, and (2) *AnyService* (0=No, 1=Yes) if the patient received any of the above identified services from the physician during their office visit.

But before we get there, we also create a summary variable for Cervical Cancer screen as patients who got PAP smear or HPVDNA testing can be combined under Cervical Cancer Screening. We name this variable *CervicalCaScreen*

```{r}
comb201618_namcs$CervicalCaScreen <- if_else(comb201618_namcs$PAP == 1 | comb201618_namcs$HPVDNA == 1, 1, 0)

```

```{r}
attach(comb201618_namcs) #attaching the dataset in use to avoid repeating the name for each of the variable below. Note that we still need to specify the dataset name for variables created after 'attaching'them to the current session
comb201618_namcs$TotalServices <- EKG + DME + COLON + AUDIO + CRYO + SPIRO + MENTAL + PT + PSYCHOTH + WOUND + ETOH + CervicalCaScreen + DEPRESS + FOOT + NEURO + PELVIC + RECTAL + RETINAL + SKIN +   SUBST + DIAEDUC + DIETNUTR + EXERCISE + GRWTHDEV + INJPREV + STRESMGT + SUBSTED + TOBACED + WTREDUC + OTHSERV + HEPTEST + SUBST + ETOHED + HIV
comb201618_namcs$TotalServices_0_1 <- ifelse(comb201618_namcs$TotalServices == 0, 0, 1)
```

```{r}
comb201618_namcs$AnyService <- DME + COLON + SIGMOID +   MENTAL + PT + PSYCHOTH + OCCUPY + CervicalCaScreen + DEPRESS + DIETNUTR + EXERCISE + STRESMGT + MAMMO + HOMEHLTH + TOBACED + CAM + BONEDENS + GENETIC +   RADTHER + HEPTEST + SUBST + ETOHED + HIV  
comb201618_namcs$AnyService_0_1 <- ifelse(comb201618_namcs$AnyService == 0, 0, 1)
detach(comb201618_namcs)
```


Now, coming to our first table (or *Table One*), we use the *TableOne* package in R to help us get a visually appealing table
We first load the required packages

```{r echo=T, results='hide', quietly=T}
source("C:/Users/Abhi/Documents/NAMCS/NAMCS_Analysis_06_12/customtab.R")
library(survey)
library(tableone)
library(tibble)
library(stringr)
library(pillar)
library(flextable)
library(tidyr)
library(dplyr)
library(officer)
library(forcats)
```

Here we are considering the variables which we include in the Table 1
```{r}
variables <- c('SEX','RACERETH', 'PAYTYPER', 'CANCER',
               'AnyService_0_1', 
               'CSTRATM','CPSUM', 'PATWT')

```
Here, we explicitly convert the binary or categorical variables into factors
```{r}
d201618_table1 <- comb201618_namcs %>% select(all_of(variables)) 

d201618_table2 <- d201618_table1 %>% mutate(SEX = as_factor(SEX),
                                           RACERETH = as_factor(RACERETH),
                                           CANCER = as_factor(CANCER),
                                           PAYTYPER = as_factor(PAYTYPER),
                                           AnyService_0_1 = as_factor(AnyService_0_1)
)

```
We will recode the *SEX* variable to help as will allow more flexibility while making the tables
```{r}
d201618_table2$SEX <- d201618_table2$SEX %>%
  fct_recode(Male = '2', Female = '1')

```
The following lines of code prepare the dataset for creating the table 1 containing the descriptive statistics
```{r}
# All variables excluding the group, survey variables AND Major (Stratification) variable as recoded above
myVars <- d201618_table2 %>% select(c(-CANCER, -CSTRATM,-PATWT,-CPSUM)) %>% names()

# All categorical variables
catVars <- d201618_table2 %>% select(where(is.factor)) %>%
  dplyr::select(-CANCER) %>% names()

svy_d201618 <- svydesign(id=~CSTRATM+CPSUM, fpc =NULL, weights = d201618_table2$PATWT, data = d201618_table2)

```

Now, with our dataset and the survey design ready, we ask to create Table 1 and use the *kableone* function to view it within the Markdown document
```{r}
tab1 <- svyCreateTableOne(vars = myVars,data = svy_d201618 ,
                          factorVars = catVars,
                          strata = "CANCER",
                          addOverall = T,
                          smd = TRUE,
                          includeNA = FALSE)
kableone(tab1)
```



```{r}

```

```{r}

```

```{r}

```

```{r}

```
